The Nuclear Forensics Problem and Statistical Methods: \\ Evaluating Machine Learning for Spent Fuel Prediction

### 2017-07-26 paper.tex
wed afternoon updates

\todo{intro, alg math, etc}

@@

\todo{intro, alg math, etc}

@@

The study does evaluate the impact of randomly introduced error of varying
amounts on the ability of the algorithms to correctly predict the burnup. It
also varies the amount of nuclides given as training to the algorithms (a full
list, a short list determined by PCA, and fission products only). However,
machine learning algorithms are heavily dependent on the inputs and parameters
given to them, such as training set sizes, learning rates, etc. Additionally,
it does not consider more advanced aglorithms. This paper adds in the
comparison of the simpler algorithms to neural nets. It also presents learning
curves for each algorithm to investigate the effect of the training set size on
prediction error. For the classification of reactor type, ROC curves are also
presented to evaluate prediction and generalization strength. For the
regression of burnup, ??? pairwise t test?.  Thus, I will first investigate
these algorithms with no error introduced, but will extend the validation from
a predetermined test set to be randomly partitioned cross validation.  And the
next study will introduce error by limiting the nuclides to only those that can
be measured with a gamma spectrometer (future work).

@@

Using predetermined test set:

1. Learning curves: For a given (randomly chosen) training set size for each
algorithm, run several trainings and predictions. This helps determine if we
are over or under training and each algorithm's robustness to over/under
fitting.

2. Regression Training Error: Confidence intervals on predictions to understand
true error versus sample error Test set must be > 30 instances, Can easily
calculate N\% confidence interval.

Using cross-validation instead:

1. redo learning curve study without having to run several trainings and
predictions for each set (bc that's already done!)

2. redo confidence interval study

>>>>

\todo{intro, alg math, explain distance metrics}

@@

\todo{intro, alg math, explain regularization}

@@

This study evaluates the impact of randomly introduced error of varying amounts
on the ability of the algorithms to correctly predict the burnup. Thus, first
investigated are the three algorithms with no error introduced, and next with
the random errors applied uniformly to each nuclide vector. Since error in a
nuclide vector is not random, in fact it is systematic and dependent on a
number of known sources of uncertainty, the next study will introduce error by
limiting the nuclides to only those that can be measured with a gamma
spectrometer (future work).

@@

Additionally, machine learning algorithms are heavily dependent on the inputs
and parameters given to them, such as training set sizes, learning rates,
regularization, etc. To evaluate the performance or tweak the model from an
algorithm, diagnostic plots will be used. Learning and validation curves will
indicate how the models are performing, initially both with respect to the
testing error and the cross validation error. As previously mentioned, these
two errors are to be compared to the training error to understand the
prediction and generalization strength with respect to training set size and
the algorithm parameters governing model complexity.

The learning curves were obtained as follows. For a given (randomly chosen)
training set size between 15 and 100\% of the total data set, several training
and prediction rounds were performed. The repetition for obtaining the testing
error is the same value as the \textit{k} in k-fold cross validation. The
testing error scenario averages the values of the obtained errors whereas the
k-fold cross-validation performs this automatically.  The validation curves
were obtained as follows. For a given parameter, the value of the parameter is
varied and \textit{k} training and prediction phases are completed, and their
errors averaged. Again, for k-fold cross-validation, these errors are already
averaged. The learning curves help determine if we are over- or under-training.
The validation curves help determine the optimal way to be robust to over- and
under-fitting.

% confidence interval stuff for future paper
%2. Regression Training Error: Confidence intervals on predictions to understand
%true error versus sample error Test set must be > 30 instances, Can easily
%calculate N\% confidence interval.
%2. redo confidence interval study
